Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-11 paper/paper.tex
update

We propose to view the computation of all these properties from an abstract algebra perspective and use a unified framework,
generic tensor-network, to solve these seeming unrelated properties all.
%We map the computation of these properties into generic tensor network contraction with specially designed tensor element algebra.
Tensor network is also known as the sum product network or einsum in different fields,
where its tensor element types are typically restricted to standard number types such as real numbers and complex numbers.
We extend this concept to generic tensor network by generalizing the tensor element type to any type being a commutative semiring.
%Traditional algorithms for finding the MIS size include the branching algorithms~\cite{Tarjan1977, Robson1986} and dynamic programming based algorithms~\cite{Courcelle1990, Fomin2013}.
Given a good contraction order, the time complexity of contracting a tensor network is $2^{{\rm tw}(G)}$~\cite{Markov2008}, where ${\rm tw}(G)$ is the treewidth of graph $G$.
Recent progress in simulating quantum circuits with random tensor networks~\cite{Gray2021, Pan2021, Kalachev2021} enables us to find such a good contraction order for a tensor network with thousands of tensors in a reasonable time.
Although this time complexity is just on par with the dynamic programming algorithms~\cite{Courcelle1990, Fomin2013} for solving the maximum size and one of the best set, tensor network has richer algebraic structure, and this structure is crutial for property computation.
%For sparse graphs, the treewidth is usually much smaller than the number of vertices~\cite{Fomin2006}. Our algorithms, on the other hand, are much more versatile than traditional methods and can be used to compute many other properties than just the MIS size.

@@

\section{Performance benchmarks}\label{sec:benchmark}
We run a single thread benchmark on CPU Intel(R) Xeon(R) CPU E5-2686 v4 @ 2.30GHz,
and its CUDA version on a GPU Tesla V100-SXM2 16G.
The results are summarized in Figure~\ref{fig:benchmark}.
The graphs that we use in benchmarks are random three regular graphs,
 a typical type of sparse graphs that has a small treewidth that asymptotically smaller than $|V|/6$~\cite{Fomin2006}.

>>>>

We propose to view the computation of all these properties from an abstract algebra perspective and use a unified framework,
generic tensor-network, to solve these seeming unrelated properties all.
%We map the computation of these properties into generic tensor network contraction with specially designed tensor element algebra.
Tensor network is also known as the sum product network or einsum in different fields,
where its tensor element types are typically restricted to standard number types such as real numbers and complex numbers.
We extend this concept to generic tensor network by generalizing the tensor element type to any type being a commutative semiring.
Given a good contraction order, the time complexity of contracting a tensor network is $2^{{\rm tw}(G)}$~\cite{Markov2008}, where ${\rm tw}(G)$ is the treewidth of graph $G$.
Recent progress in simulating quantum circuits with random tensor networks~\cite{Gray2021, Pan2021, Kalachev2021} enables us to find such a good contraction order for a tensor network with thousands of tensors in a reasonable time.
Although this time complexity is just on par with the dynamic programming algorithms~\cite{Courcelle1990, Fomin2013} for solving the maximum size and one of the best set, tensor network has richer algebraic structure, and this structure is crutial for property computation.
%For sparse graphs, the treewidth is usually much smaller than the number of vertices~\cite{Fomin2006}. Our algorithms, on the other hand, are much more versatile than traditional methods and can be used to compute many other properties than just the MIS size.

@@

\section{Performance benchmarks}\label{sec:benchmark}
We run a single thread benchmark on CPU Intel(R) Xeon(R) CPU E5-2686 v4 @ 2.30GHz,
and its CUDA version on a GPU Tesla V100-SXM2 16G.
The results are summarized in Figure~\ref{fig:benchmark}.
The graphs that we use in benchmarks are random three regular graphs,
 a typical type of sparse graphs that has a small treewidth that asymptotically smaller than $|V|/6$~\cite{Fomin2006}.
In this benchmark, we do not include traditional algorithms for finding the MIS size like the branching algorithms~\cite{Tarjan1977, Robson1986} and dynamic programming based algorithms~\cite{Courcelle1990, Fomin2013},
because to the best of our knowledge, most of the properties mentioned in this paper are not computable within these frameworks.
